---
title: "Lab 01 - Hello R"
author: "Cynthia Deng"
date: "01-18-2026"
output: 
  html_document: 
    fig_height: 4
    fig_width: 9
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
data("datasaurus_dozen")
```

## Exercises

### Exercise 1
```{r}
data("datasaurus_dozen")

datasaurus_dozen %>%
  count(dataset) %>%
  print()
glimpse(datasaurus_dozen)

```

The datasaurus_dozen file has 1846 rows and 3 columns. The three variables are dataset, x, and y.

### Exercise 2

The answers for this exercise are already provided to help guide you. Clean up the narrative to reflect what you want to submit. Let’s break it into steps:

First, let's plot the data in the dino dataset:

```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")

ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

Next, let's calculate the correlation between `x` and `y` in this dataset:

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```
1. The datasaurus_dozen dataframe that contains 13 datasets, we first use filter to find the rows where dataset = 'dino_data' and saved them in a new data frame called dino_data.
2. We used ggplot to create a scatterplot of the x and y variabels from the dino_data data frame.
3. We used cor (x, y) to calculate the Pearson correlation between x and y variables in dino_data. We assigned the correlation we calculated to a variable named r and presented the correlation in a tibble.


### Exercise 3

1. Filter dataset = 'star' from the large datasaurus_dozen data frame. Save the filtered result to a new data frame called star_data. code adapted from the {r plot-dino} code chunk.
2. Use ggplot to create a scatterplot of the x and y variabels from the star_data data frame

```{r plot-star}
star_data <- datasaurus_dozen %>%
  filter(dataset == "star")

ggplot(data = star_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

calculate correlation for star_data. Code adapted from sample code for dino_data

```{r cor-star}
star_data %>%
  summarize(r = cor(x, y))
```
The correlation for the star dataset is −0.0629611, and the correlation for the dino dataset is −0.06447185. Both correlations are very weak (close to 0) . But the correlation for dino is slightly stronger than correlation for star.


### Exercise 4
1. edited the name of the code chunk to plot-circle and cor-circle to make it easier to see the purpose of each code chunk.
2. Filter dataset = 'circle' from the large datasaurus_dozen data frame. Save the filtered result to a new data frame called star_data. code adapted from the {r plot-dino} code chunk.
3. Use ggplot to create a scatterplot of the x and y variabels from the circle_data data frame.

```{r plot-circle}
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")

ggplot(data = circle_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

```{r cor-circle}
circle_data %>%
  summarize(r = cor(x, y))
```
4. calculate correlation for circle_data. Code adapted from sample code for dino_data.
The correlation for the circle dataset is -0.06834336	, and the correlation for the dino dataset is −0.06447185. The correlation for dino is slightly weaker than correlation for circle.

### Exercise 5

This is your final exercise! Add code and narrative as needed. Be creative and show off what you’ve learned so far.
Add code and narrative as needed.

To add additional R chunks, you can either type the backticks, curly braces, and the letter `r`, or use the "Insert Chunk" button above (the green C+ icon). Have fun!
```{r}
ggplot(datasaurus_dozen, aes(x = x, y = y, color = dataset)) +
  geom_point() +
  facet_wrap(~dataset, ncol = 3) +
  theme(legend.position = "none")
```
1.We used ggplot to create scatterplots of x and y for all datasets in datasaurus_dozen. The data are faceted by dataset, so that each dataset appears in its own panel.
```{r}
datasaurus_dozen %>%
  group_by(dataset) %>%
  summarize(r = cor(x, y)) %>%
  print(13, width = Inf)
```
2. We calculate the Pearson correlation between x and y for each dataset. We used groupby to group the data by the dataset variable. As we can see the printed tibble, the correlations are very similar across the 13 different data frames, although their plots were very different.

